---
title: "project3"
author: "group1"
date: "March 16, 2018"
output:
  html_document:
    theme: cerulean
    hightlight: tango
    css: styles.css
---

```{r setup, include=FALSE}
library(plyr)
library(tidyverse)
library(splitstackshape)
library(magrittr)
library(knitr)
```

## Our Project {.tabset}

```{r, warning=FALSE, message=FALSE}
raw.data <- read_csv('https://raw.githubusercontent.com/brian-cuny/607project3/master/multipleChoiceResponses.csv', na=c('', 'NA')) %>%
  subset(DataScienceIdentitySelect == 'Yes' & CodeWriter == 'Yes') %>%
  rowid_to_column('id')
```

### Profile of a Data Scientist: Justin



### Learning Platform Usefulness: Hovig



### Learning Categories: Brian

This subset of data examines how data scientists learned their core skill set. Each data scientist was asked to assign each category a percent from 0 to 100 indicating how much of their education was made up of this source.

First, a list of categories was extrated and formatted.

```{r}
tidy.names <- names(raw.data)[61:66] %>% 
  str_extract('(?<=LearningCategory)(\\w+)') %>% 
  str_replace_all('(?<=[a-z])([A-Z])', '_\\1') %>% 
  tolower()
tidy.names %>% kable()
```

The data was tidied and the categories were converted to factors, to aid in analysis.

```{r}
tidy.data <- raw.data %>%
  select(c(1, 61:66)) %>%
  setNames(c('id', tidy.names)) %>%
  gather('category', 'percent', 2:7, na.rm=TRUE)

tidy.data$percent %<>% as.numeric()

tidy.data$category %<>% factor(levels=tidy.names, ordered=TRUE)
tidy.data %>% head(10) %>% kable()
```

Summary statistics tell an intersting story. No source averaged more than 50% of the sets learning sources. This would seem to indicate that data scientists learn from a diverse set of sources. The 'other' category's mean is nearly 0 also indicating that the other categories account for nearly all learning sources. 

```{r}
tidy.summary.data <- tidy.data %>% 
  group_by(category) %>% 
  summarise(avg=mean(percent), sd=sd(percent))
tidy.summary.data %>% kable()
```

The boxplots support the summary statistics. Each category has numerous upper ourliers indicating that data scientists who learned most or entirely from one source were rare.

```{r}
ggplot(tidy.data) +
  geom_boxplot(aes(category, percent)) +
  xlim(tidy.names %>% rev()) +
  coord_flip() + 
  labs(x='Learning Source', 
       y='Proportion',
       title='Data Scientists Learn From Diverse Sources'
  )
```

The final more clearly shows the diversity in learning styles. This indicates that not only do data scientists learn from a variety of sources, but every data scientist's sources vary in importance. This highlights the idea that there is not right or wrong way to learn to become a data scientist. At the same time, as the four major categories amount for nearly 100% of education, this means that there are no "secret" learning sources.

```{r}
ggplot(tidy.data) +
  geom_bar(aes(category, fill=percent %>% 
                                round_any(10) %>% 
                                factor(seq(0, 100, 10))
              ), position=position_fill(reverse=TRUE)
          ) +
  scale_color_brewer(palette='Set1') + 
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  labs(x='Learning Source', 
       y='Proportion',
       title='Data Scientists Learn From Diverse Sources',
       fill='Percent'
  )
```

### Common Job Algorithms: Rose


### Work Tools Freqeuncy: Meaghan


### Work Challenges: Albert


### Conclusion (writing to csv/importing to sql?)
