---
title: "test2"
author: "Meaghan Burke"
date: "March 24, 2018"
output: html_document
---
###Data Scientists Work Tools:

In this section, we determine which tools are the most frequently used by a data scientist. Data scientists need tools that will perform data analysis, data warehousing, data visualization and machine learning. We suspect that a typical data scientist uses a multitude of tools to satisfy these components. 
```{r results='hide', message=FALSE, warning=FALSE}
library(plyr)
library(tidyverse)
library(splitstackshape)
library(magrittr)
library(stringr)
```

```{r results='hide', message=FALSE, warning=FALSE}
raw.data <- read_csv("https://raw.githubusercontent.com/brian-cuny/607project3/master/multipleChoiceResponses.csv", na = c('')) %>%
  subset(DataScienceIdentitySelect == 'Yes' & CodeWriter == 'Yes') %>%
  rowid_to_column('id')

raw.data <- as.data.table(raw.data)
```

```{r}
tidy.names <- names(raw.data)[83:132]%>% 
  str_extract('(?<=WorkToolsFrequency)(\\w+)') %>% 
  str_replace_all('(?<=[a-z])([A-Z])', '_\\1') 

tools.data <- raw.data %>%
  select(c(1, 82:84)) %>%
  setNames(c('id', 'tool_used', "temp_1", "temp_2"))%>%
  unite_("tool_used", c("tool_used","temp_1","temp_2"))%>%
  mutate(tool_used = (str_replace_all(tool_used, '/', ',')),
         tool_used = (str_replace_all(tool_used, '_', ',')))%>%
  mutate(tool_counter =1)
tools.data <- cSplit(tools.data, 'tool_used', ',')
```



```{r}
id.tool.df <- tools.data %>%
  gather(tool_group, tool, names(tools.data)[3:63])%>%
  group_by(id, tool)%>%
  summarise(sum_tool = sum(tool_counter))%>%
  drop_na()%>%
  filter(!tool %in% c("Rarely", "Often",
                      "Sometimes", "Most of the time"))
```

```{r}
summary.tool.df <- tools.data %>%
  gather(tool_group, tool, names(tools.data)[3:63])%>%
  group_by(tool)%>%
  summarise(sum_tool = sum(tool_counter))%>%
  drop_na()%>%
  arrange(desc(sum_tool))%>%
  filter(!tool %in% c("Rarely", "Often",
                      "Sometimes", "Most of the time"))%>%
  mutate(percent_total = round((sum_tool/ sum(sum_tool))*100,digits = 2))

```

```{r}
ggplot(head(summary.tool.df,15), aes(x=reorder(tool, -sum_tool), y=percent_total)) + 
  geom_bar(stat="identity", width=.5, fill="tomato3") +
  geom_text(aes(label=percent_total))+
  labs(x='Tool', 
       y='Percent Total',
       title="Top 15 Data Science Tools", 
       caption="Source: Multiple Choice Responses") + 
  theme(axis.text.x = element_text(angle=65, vjust=0.6))
```

```{r}
frequency.data <- raw.data %>%
  select(c(1, 83:132)) %>%
  setNames(c('id', tidy.names))

id.frquency.table <-frequency.data %>% 
  gather(tool_name, frequency_id, names(frequency.data)[2:51])%>%
  filter(frequency_id %in% c("Rarely", "Often",
                             "Sometimes", "Most of the time"))%>%
  arrange(id)
```
```{r}
summary.frquency.table <- frequency.data %>% 
  gather(tool_name, frequency_id, names(frequency.data)[2:51])%>%
  filter(frequency_id %in% c("Rarely", "Often",
                             "Sometimes", "Most of the time"))%>%
  mutate(freq_counter =1) %>%
  group_by(tool_name,frequency_id)%>%
  summarise(sum_feq = sum(freq_counter))%>%
  arrange(desc(sum_feq))
```

```{r}
ordering <- c('Most of the time', 'Often', 'Sometimes', 'Rarely')

ggplot(head(summary.frquency.table,50), aes(x = frequency_id, y = sum_feq, fill = tool_name)) + 
  geom_bar(stat = "identity") + 
  facet_wrap(~tool_name) + 
  ylab("Number of times a response was selected") + 
  xlab('') + 
  xlim(ordering) +
  theme(legend.position="none") +
  theme(axis.text.x = element_text(angle = 90, 
                                   vjust = 0.5, 
                                   hjust = 1))

```
The survey reports a wide variety of responded tools, all of which seemingly address the different needs of a data scientist. Among the top 15 tools, there were a healthy mix of analysis tools, warehousing tools and visualization tools. This indicates that data scientists choose an assortment of tools when addressing tasks. It also indicates a potential overlap in the functionality and features of the tools available. It will be interesting to see how the landscape of tools will evolve as the field of data science matures.

